Meeting --- Summary LLM

### 2024-05-28 meetings/xx-xx-2024.qmd
improvements on gradients

- Gradient spikes first, then large step in decoder / encoder
- These steps seem to affect one hidden unit more strongly than the other
- Increase of Decoder Weight gradients seems to precede the increase of the -LL
- The other gradients seem to react to this jump
- Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...

@@

![](./figures/dashboard_last10_540.png)

@@

![](./figures/dashboard_last10_160.png)

@@

![](./figures/dashboard_last10_350.png)

>>>>

- Gradient spikes first, then large step in decoder / encoder
- The other gradients seem to react to this jump
- Gradient $\to$ Step $\to$ Other gradients $\to$ Step $\to$ -LL jumps 
- [`torch.nn.utils.clip_grad_norm_(ae.parameters(), max_norm=dMaxGrad)`](https://pytorch.org/docs/stable/generated/torch.nn.utils.clip_grad_norm_.html#torch.nn.utils.clip_grad_norm_) seems to help. This clips the gradient if the norm exceeds `dMaxGrad`. [See here](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/readings/L15%20Exploding%20and%20Vanishing%20Gradients.pdf#page=6.78).

@@

![More spikes I](./figures/dashboard_last10_540.png)

@@

![More spikes II](./figures/dashboard_last10_160.png)

@@

![More spikes III](./figures/dashboard_last10_350.png)

Optimization with gradient-clipping at $dMaxGrad = 0.3$.

![](./figures/dashboard_clipping_fixed.png)

Optimization with gradient-clipping at $dMaxGrad = 0.1$. Deeper in the optimization those spikes seem to become a problem again, at a lower scale. Oscill. of gradients later in the optimization.

![Clipping at 0.1](./figures/dashboard_clip_01.png)

Optimization with gradient-clipping at $dMaxGrad = 0.05$. This model also has a normalization layer after the hidden layer.

![Clipping at 0.05 and normalization layer](../mlartifacts/437740661089244805/50054b1717e444eeb1d4acd408e1e77d/artifacts/plots/dashboard.png)
